Pathogen Spillover - dissertation chapter

### 2018-09-28 pathogen-spillover-diss-ch.Rmd
Revisit modeling and make model output figures.

---
title: "Pathogen Spillover - dissertation chapter"
author: "Whalen Dillon"
date: "July 18, 2016"
output:
  html_document:
    self_contained: no
    toc: yes
  pdf_document: default
  word_document: default
---

@@

![Conceptual Path Model](figures/temperature-path-model-updated.png)

@@

```{r load data, echo=F, message=F}
library(knitr)
load("climate_path_model_data.RData")
library(piecewiseSEM)# version 1.1.1, tried most recent version & kept getting errror 7/18/2016 b/c matrices rows were unequal in sem.missing.paths function
## Avoided this error in 1.1.2 by removing observations with NA values, which gives the same results as before, so maybe it wasn't properly dropping missing values when extracting info from the models.
library(plyr); library(dplyr, verbose = F)
library(lme4, verbose = F); library(nlme, verbose = F)
# library(lmerTest, verbose = F)

@@

## make consolidated data frame only of variables in the path model
oakplots.model.data <- select(oakplots.sub, plotid, sample_year, oak.inf, tot_lfct, tot_lfct.log, hrs_abv25ds, avg_hrs1422_wet_t_t1, avg_hrs1422_wet_t_t1.log, umca_density_ha, umca_density_ha.log, H.2005, twi15m, twi15m.log#, twi15m.log.scl, tot_lfct.log.scl, hrs_abv25ds.scl, avg_hrs1422_wet_t_t1.log.scl, umca_density_ha.log.scl, shannons2005.scl
                              )

@@

oakplots.model.data.sub <- oakplots.model.data %>% 
      filter(plotid!="hood03", plotid!="hood05", plotid!="jlsp04", plotid!="johns02", plotid!="sugar17")

write.csv(oakplots.model.data.sub, "pathogen_spillover_data.csv", row.names = F)
```

@@

```{r hours >25 path model all oak plots unscaled, echo=F}
# slc_nlme <- lme(tot_lfct.log ~ umca_density_ha.log + hrs_abv25ds + avg_hrs1422_wet_t_t1.log + H.2005, random = ~1|plotid, data = oakplots.model.data)

model_list1 <- list(
      shannons <- lm(
            H.2005 ~ twi15m, 
            data = filter(oakplots.model.data, sample_year == 2006)),

@@

      oak_infection <- glmer(
            oak.inf ~ tot_lfct.log + avg_hrs1422_wet_t_t1.log + H.2005 + (1|plotid) + (1|sample_year), family = binomial(link = "logit"), 
            data = oakplots.model.data)
)

# refit model with missing path from diversity to umca density
model_list1_refit <- list(
      shannons_refit <- lm(
            H.2005 ~ twi15m, 
            data = filter(oakplots.model.data, sample_year == 2006)),

      umca_density_refit <- lmer(
            umca_density_ha.log ~ H.2005 + twi15m + (1|plotid) + (1|sample_year), 
            data = oakplots.model.data),

      slc_refit <- lmer(
            tot_lfct.log ~ umca_density_ha.log + hrs_abv25ds + avg_hrs1422_wet_t_t1.log + H.2005 + (1|plotid) + (1|sample_year), 
                  data = oakplots.model.data),

      wet_hours_1422_refit <- lmer(
            avg_hrs1422_wet_t_t1.log ~ twi15m + (1|plotid) + (1|sample_year), 
                             data = oakplots.model.data),

      hrs_abv25_refit <- lmer(
            hrs_abv25ds ~ twi15m + (1|plotid) + (1|sample_year), 
            data = oakplots.model.data),

      oak_infection_refit <- glmer(
            oak.inf ~ tot_lfct.log + avg_hrs1422_wet_t_t1.log + H.2005 + (1|plotid) + (1|sample_year), 
            family = binomial(link = "logit"), 
            data = oakplots.model.data)
)
```

```{r path-model-no-lide-plots, echo=F}
model_list1.noLIDE <- list(
      shannons <- lm(H.2005 ~ twi15m, data = filter(oakplots.model.data.sub, sample_year == 2006)),

@@

**Residual plots of Shannon's Diversity Index model**
```{r shannons residuals plot, echo=F}
plot(shannons, main = "OLS regression of diversity ~ wetness index residuals")
plot(shannons$model$twi15m, resid(shannons),
     xlab = "wetness index",
     ylab = "model residuals",
     main = "Diversity model residuals vs. wetness index predictor")
```

**Residual plots of bay laurel density model**
```{r bay laurel density residuals plot, echo=F}
plot(umca_density, main = "LMM of bay laurel density residuals")
plot(umca_density@frame$twi15m, resid(umca_density),
     xlab = "wetness index",
     ylab = "model residuals",
     main = "Bay laurel density model residuals vs. wetness index predictor")
```

**Residual plots of symptomatic leaf count model**
```{r symptomatic leaf count residuals plot, echo=F}
plot(resid(slc), main = "LMM of symptomatic leaf count residuals")
# qqnorm(resid(slc), main="Q-Q plot for conditional residuals - SLC")
# qqnorm(ranef(slc)$sample_year$`(Intercept)`, 
#        main="Q-Q plot for Sample Year")
# qqnorm(ranef(slc)$plotid$`(Intercept)`, 
#        main="Q-Q plot for plot ID")

plot(slc@frame$umca_density_ha.log, resid(slc),
     xlab = "bay laurel density",
     ylab = "model residuals",
     main = "SLC model residuals vs. bay laurel density predictor")
plot(slc@frame$hrs_abv25ds, resid(slc),
     xlab = "# dry-season hours >25 C",
     ylab = "model residuals",
     main = "SLC model residuals vs. dry-season hours >25 C predictor")
plot(slc@frame$avg_hrs1422_wet_t_t1.log,
     xlab = "avg. wet-hours 14-22C",
     ylab = "model residuals",
     main = "SLC model residuals vs. wet-hours predictor")
plot(slc@frame$H.2005, resid(slc),
     xlab = "Shannon's diversity index",
     ylab = "model residuals",
     main = "SLC model residuals vs. Shannon's diversity predictor")
# plot(slc@frame$twi15m.log.scl, resid(slc),
#      xlab = "wetness index",
#      ylab = "model residuals",
#      main = "SLC model residuals vs. wetness index predictor")
```
```{r pairs plot of slc predictor variables, echo=F}
pairs(na.omit(select(oakplots.sub, umca_density_ha.log, hrs_abv25ds, avg_hrs1422_wet_t_t1.log, H.2005, twi15m)),
      lower.panel = panel.cor, diag.panel = panel.hist,
      main = "Predictor variables for symptomatic leaf count models")
```

**Residual plots of average number of wet-hours model**
```{r wet-hours residuals plot, echo=F}
plot(wet_hours_1422, main = "LMM of avg. # of hours 14-22 C current & prev. year residuals")
# plot(wet_hours_1422@frame$umca_density_ha.log.scl, resid(wet_hours_1422),
#      xlab = "bay laurel density",
#      ylab = "model residuals",
#      main = "Wet-hours model residuals vs. bay laurel density predictor")
plot(wet_hours_1422@frame$twi15m, resid(wet_hours_1422),
     xlab = "bay laurel density",
     ylab = "model residuals",
     main = "Wet-hours model residuals vs. bay laurel density predictor")
```

**Residual plots of dry-season temperature >25C model**
```{r dry-season hours residuals plot, echo=F}
plot(hrs_abv25, main = "LMM of # of hours >25 C during prev. dry season residuals")
plot(hrs_abv25@frame$twi15m, resid(hrs_abv25),
     xlab = "wetness index",
     ylab = "model residuals",
     main = "Dry-season hours model residuals vs. wetness index predictor")
# plot(hrs_abv25@frame$umca_density_ha.log.scl, resid(hrs_abv25),
#      xlab = "wetness index",
#      ylab = "model residuals",
#      main = "Dry-season hours model residuals vs. wetness index predictor")
```

**Residual plots of oak infection model**
```{r oak infection residuals plot, echo=F}
plot(oak_infection, main = "Binomial GLMM of oak infection residuals")
plot(oak_infection@frame$tot_lfct.log, resid(oak_infection),
     xlab = "symptomatic leaf count",
     ylab = "model residuals",
     main = "Oak infection model residuals vs. symptomatic leaf count predictor")
plot(oak_infection@frame$avg_hrs1422_wet_t_t1.log, resid(oak_infection),
     xlab = "avg. wet-hours 14-22C",
     ylab = "model residuals",
     main = "Oak infection model residuals vs. wet-hours predictor")
plot(oak_infection@frame$H.2005, resid(oak_infection),
     xlab = "Shannon's diversity index",
     ylab = "model residuals",
     main = "Oak infection model residuals vs. diversity predictor")
```
```{r pairs plot of oak infection predictor variables, echo=F}
pairs(na.omit(select(oakplots.sub, tot_lfct.log, avg_hrs1422_wet_t_t1.log, H.2005)),
      lower.panel = panel.cor, diag.panel = panel.hist,
      main = "Predictor variables for oak infection model")
```

@@

```{r tables of path coefficients, echo=F}
kable(raw_coefficients, digits = 4,
      caption = "Raw path coefficients")
# kable(raw_coefficients_noUMCA,
#       caption = "Raw path coefficients - no UMCA density")

>>>>

---
title: "Pathogen Spillover - dissertation chapter"
author: "Whalen Dillon"
date: "`r Sys.Date()"
output:
  html_document:
    self_contained: no
    toc: true
    toc_float: true
  pdf_document: default
  word_document: default
---

@@

![Conceptual Path Model](figures/Fig2-conceptual-path-model-updated.png)

@@

```{r load data, echo=F, message=F}
library(knitr)
load("climate_path_model_data.RData")

# devtools::install_github("jslefche/piecewiseSEM@devel", build_vignette = TRUE)
library(piecewiseSEM)# version 1.1.1, tried most recent version & kept getting errror 7/18/2016 b/c matrices rows were unequal in sem.missing.paths function
## Avoided this error in 1.1.2 by removing observations with NA values, which gives the same results as before, so maybe it wasn't properly dropping missing values when extracting info from the models.
library(plyr); library(dplyr, verbose = F)
library(lme4, verbose = F); library(nlme, verbose = F)
library(ggplot2)
library(viridisLite); library(viridisLite)
# library(lmerTest, verbose = F)

@@

## make consolidated data frame only of variables in the path model
oakplots.model.data <- select(oakplots.sub, plotid, sample_year, oak.inf, inf_oak_ct, uninf_oak_ct, tot_bay, umca_density_ha, tot_lfct, tot_lfct.log, hrs_abv25ds, avg_hrs1422_wet_t_t1, avg_hrs1422_wet_t_t1.log, umca_density_ha.log, H.2005, twi15m, twi15m.log#, twi15m.log.scl, tot_lfct.log.scl, hrs_abv25ds.scl, avg_hrs1422_wet_t_t1.log.scl, umca_density_ha.log.scl, shannons2005.scl
                              )

@@

# Exclude LIDE plots
oakplots.model.data.sub <- oakplots.model.data %>% 
      filter(plotid!="hood03", plotid!="hood05", plotid!="jlsp04", plotid!="johns02", plotid!="sugar17")

# write.csv(oakplots.model.data.sub, "pathogen_spillover_data.csv", row.names = F)

@@

```{r hours >25 path model all oak plots unscaled, echo=F, eval=F}
# slc_nlme <- lme(tot_lfct.log ~ umca_density_ha.log + hrs_abv25ds + avg_hrs1422_wet_t_t1.log + H.2005, random = ~1|plotid, data = oakplots.model.data)

model_list1 <- list(
      shannons <- lm(
            H.2005 ~ twi15m, 
            data = filter(oakplots.model.data, sample_year==2006)),

@@

      oak_infection <- glmer(
            oak.inf ~ tot_lfct.log + avg_hrs1422_wet_t_t1.log + H.2005 + (1|plotid) + (1|sample_year), family = binomial(link = "logit"), 
            data = oakplots.model.data)
)
```

Add some re-calcualted variables and explore the bay density~diversity relationship.
```{r re-examine data}

# refit model with missing path from diversity to umca density
oaks.data <- oakplots.model.data %>% 
      rename(hrs25 = hrs_abv25ds, 
             hrs1422 = avg_hrs1422_wet_t_t1.log, 
             lfct = tot_lfct.log, 
             umcadens = umca_density_ha.log, 
             h2005 = H.2005, 
             oakinf = oak.inf, 
             year = sample_year)

oaks.data <- oaks.data %>% 
      mutate(tot_bay.ln = log(tot_bay),
             tot_bay.ln1 = log1p(tot_bay),
             tot_bay0.1 = ifelse(tot_bay==0, 0.1, tot_bay),
             tot_bay.ln.1 = log(tot_bay0.1)
             )

GGally::ggpairs(select(oaks.data, h2005, twi15m, tot_bay, tot_bay.ln, tot_bay.ln1))

ggplot(oaks.data, 
       aes(h2005, tot_bay)) +
      geom_point() +
      geom_smooth(method = "lm") +
      scale_fill_viridis_d() +
      scale_color_viridis_d() +
      # geom_density(alpha = .8) +
      theme_classic() +
      NULL

ggplot(oaks.data, aes(tot_bay)) +
      geom_density() +
      theme_classic()
ggplot(oaks.data, aes(tot_bay.ln)) +
      geom_density() +
      geom_density(aes(tot_bay.ln1), color = "blue") +
      theme_classic()

ggplot(oaks.data) +
      geom_density(aes(h2005)) +
      theme_classic()

oaks.data %>% 
      # filter(tot_bay>0) %>% 
      filter(year==2006) %>%
ggplot(aes(h2005, tot_bay.ln)) +
      geom_point() +
      geom_smooth(method = "lm") +
      geom_smooth(aes(y=tot_bay.ln1), method = "lm", color = "red") +
      # geom_smooth(aes(y=tot_bay.ln.001), method = "lm", color = "tan4") +
      # geom_density(alpha = .8) +
      theme_classic() +
      NULL
## DON'T ADD 1 & LOG-TRANSFORM
summary(oaks.data)
filter(oaks.data, is.infinite(tot_bay.ln))$tot_bay
oaks.data$tot_bay.ln[is.infinite(oaks.data$tot_bay.ln)] <- 0

m1 <- lmer(tot_bay ~ h2005 + twi15m + (1|plotid) + (1|year),
           data = oaks.data, na.action = na.omit)
m1a <- lmer(tot_bay.ln ~ h2005 + twi15m + (1|plotid) + (1|year),
           data = oaks.data, na.action = na.omit)
# anova(m1,m1a)
summary(m1)
plot(m1)
# plot(m1a)
# summary(m1a)

m2 <- mgcv::gamm(tot_bay ~ s(h2005) + twi15m, random=list(year=~1, plotid=~1), data = oaks.data, na.action = na.omit)
summary(m1)
summary(m2[[1]])
summary(m2[[2]])
plot(m2[[1]])
plot(m2[[2]])
mgcv::vis.gam(m2)
m3 <- gamm4::gamm4(tot_bay ~ s(h2005) + twi15m, random = ~(1|plotid)+(1|year), data = oaks.data, na.action = na.omit)
summary(m3[[1]])
plot(m3[[1]])
summary(m3[[2]])
plot(m3[[2]])

m4 <- glmer(tot_bay ~ h2005 + twi15m + (1|plotid) + (1|year),
           data = oaks.data, na.action = na.omit, family = poisson)
# par(mfrow=c(2,1))
plot(m1); plot(m4)
summary(m1)
summary(m4)
anova(m1, m4)

m5 <- glmer.nb(tot_bay ~ h2005 + twi15m + (1|plotid) + (1|year),
           data = oaks.data, na.action = na.omit)
summary(m5)
plot(m5)

dat <- filter(oaks.data, year==2006)
m6 <- glm(tot_bay ~ h2005 + twi15m, data = dat, family = quasipoisson, na.action = na.fail)
summary(m6)
# plot(m6)
(c_hat <- deviance(m6) / df.residual(m6))
library(MuMIn)
dredge(m6, rank = "QAIC", chat = c_hat)
dredge(m6, rank = "AIC")
hacked.quasipoisson <- function(...) {
    res <- quasipoisson(...)
    res$aic <- poisson(...)$aic
    res
}
QAIC(update(m6, family = hacked.quasipoisson, chat = c_hat)

# library(gamlss)
# m7 <- gamlss(tot_bay ~ h2005 + twi15m, data = oaks.data)
# summary(m7)
# plot(m7)

# GGally::ggpairs(select(oaks.data, tot_lfct,tot_bay,h2005))
m1 <- glmer(tot_lfct ~ tot_bay + hrs25 + hrs1422 + h2005 + (1|plotid) + (1|year), family = poisson, data = oaks.data, na.action = na.omit)
summary(m1)
m2 <- glmer.nb(tot_lfct ~ tot_bay + hrs25 + hrs1422 + h2005 + (1|plotid) + (1|year), data = oaks.data, na.action = na.omit)
summary(m2)
anova(m2,lfctmod)
plot(lfctmod)
```

```{r refit model with H bay density path, eval=F}
# The refit model has path between H and UMCA density ####
model_list1_refit <- list(
      shannonmod <- lm(h2005 ~ twi15m, data = filter(oaks.data, year==2006), na.action = na.omit),

      densmod <- lmer(umcadens ~ h2005 + twi15m + (1|plotid) + (1|year),
           data = oaks.data, na.action = na.omit),

      lfctmod <- lmer(lfct ~ umcadens + hrs25 + hrs1422 + h2005 + (1|plotid) + (1|year), data = oaks.data, na.action = na.omit),

      wetmod <- lmer(hrs1422 ~ twi15m + (1|plotid) + (1|year), data = oaks.data, na.action = na.omit),

      hotmod <- lmer(hrs25 ~ twi15m + (1|plotid) + (1|year), data = oaks.data, na.action = na.omit),

      # hrs25 %~~% hrs1422,

      infmod <- glmer(oakinf ~ lfct + hrs1422 + h2005 + (1|plotid) + (1|year), family = binomial(link = "logit"), data = oaks.data, na.action = na.omit)
)
summary(lmer(h2005 ~ twi15m + (1|plotid) + (1|year), data = oaks.data, na.action = na.omit))
summary(densmod)
summary(lfctmod)
summary(wetmod)
summary(hotmod)
summary(infmod)

# summary(oaks.data)

# ?psem I get errors with row matching when trying to use the new fucntion. I think it may have more to do with the new version of R and how that affects the internal sub-modeling when there are complex mixed effects models.

```{r Update model list}

# Updated SEM uses untransformed UMCA variables ####
# lfct is the ln(tot_lfct+1)
      # lmer(h2005 ~ twi15m + (1|plotid) + (1|year), data = oaks.data, na.action = na.omit),
m_list_update <- list(

            shannon.mod <- lm(h2005 ~ twi15m, data = filter(oaks.data, year==2006), na.action = na.omit),
      # changed the bay response to be log-transformed 
      umca.mod <- lmer(tot_bay.ln1 ~ twi15m + (1|plotid) + (1|year), data = oaks.data, na.action = na.omit),

      lfct.mod <- lmer(lfct ~ tot_bay.ln1 + hrs25 + hrs1422 + (1|plotid) + (1|year), data = oaks.data, na.action = na.omit),

      wet.mod <- lmer(hrs1422 ~ twi15m + (1|plotid) + (1|year), data = oaks.data, na.action = na.omit),

      hot.mod <- lmer(hrs25 ~ twi15m + (1|plotid) + (1|year), data = oaks.data, na.action = na.omit),

      # hrs25 %~~% hrs1422,

      inf.mod <- glmer(oakinf ~ lfct + hrs1422 + h2005 + (1|plotid) + (1|year), family = binomial(link = "logit"), data = oaks.data, na.action = na.omit)

)

# ggplot(oaks.data, aes(tot_bay.ln1, lfct)) +
#       geom_point() +
#       geom_smooth(method = "lm", color = "red") +
#       geom_smooth() +
#       theme_classic()

```{r path-model-no-lide-plots, echo=F, eval=F}
model_list1.noLIDE <- list(
      shannons <- lm(H.2005 ~ twi15m, data = filter(oakplots.model.data.sub, sample_year == 2006)),

@@

**Residual plots of Shannon's Diversity Index model**
```{r shannon.mod plot, echo=F}

library(sjPlot)
plot_model(shannon.mod, type = "diag", show.values = T, grid = TRUE)
plot_model(shannon.mod, type = "slope", show.values = T)
plot_model(shannon.mod, type = "resid", show.values = T)
plot_model(shannon.mod, type = "std", show.values = T, 
           title = "Standardized coefficient estimate")
S.plot <- plot_model(shannon.mod, type = "slope")

# line_color <- "black"
# S.plot$data
# str(S.plot)

(shannon.plot <- ggplot(oaks.data, aes(twi15m, h2005)) +
      geom_point(alpha = .7) +
      # geom_smooth(method = "loess", se = F, color = "gray50", linetype = "dashed") +
      geom_smooth(method = "lm", color = "black") +
      theme_classic() +
      xlab("Topographic index") +
      ylab("Tree diversity (H')") +
      theme(strip.background = element_blank(),
            strip.text = element_blank()) +
      NULL
      )

# plot(shannons, main = "OLS regression of diversity ~ wetness index residuals")
# plot(shannons$model$twi15m, resid(shannons),
#      xlab = "wetness index",
#      ylab = "model residuals",
#      main = "Diversity model residuals vs. wetness index predictor")

**Residual plots of bay laurel density model**
```{r umca.mod plot, echo=F}

plot_model(umca.mod, type = "diag")
plot_model(umca.mod, show.values = T, type = "est", 
           title = "Coefficient estimates - UMCA density")
plot_model(umca.mod, show.values = T, type = "std", 
           title = "Standardized coefficient estimates - UMCA density")
plot_model(umca.mod, show.values = T, type = "re")

# umca_vs_h2005 <- plot_model(umca.mod, show.values = T, type = "pred", terms = c("twi15m"), pred.type = "fe")
# umca_vs_h2005 +
#       theme_classic()

# oaks.data %>% 
#       filter() %>% 
#       ggplot(aes(h2005, log1p(tot_bay))) +
#       geom_point() +
#       geom_smooth(method = "lm") +
      # facet_grid(.~year)
# qplot(x=h2005, y=tot_bay, data = oaks.data)

# library(ggeffects)
# umca.mod_predict <- ggpredict(model = umca.mod.update, terms = "h2005", type = "re", full.data = T)
# ggplot(umca.mod_predict, aes(x, predicted)) +
#       # data = umca.mod.update@frame,
#       geom_point(aes(x, predicted), show.legend = F) +
#       # geom_smooth(se = F) +
#       geom_smooth(method = "lm") +
#       # facet_grid(.~year) +
#       xlab("Shannon's diversity index (H')") +
#       ylab("Bay laurel density (stems/ha)") +
#       theme_classic() +
#       NULL

# oaks.data$h2005.sq <- oaks.data$h2005^2
# oaks.data$tot_bay.log <- log1p(oaks.data$tot_bay)
# umca.mod.update <- lmer(tot_bay.log ~ h2005 + twi15m + (1 | plotid) + (1 | year), data = oaks.data)
# plot(umca.mod)
# plot_model(umca.mod.update, show.values = T, type = "resid")
# plot_model(umca.mod.update, show.values = T, type = "diag")

# umca_vs_twi15m <- plot_model(densmod, show.values = T, type = "eff", terms = c("twi15m"), pred.type = "re")

# plot(umca_density, main = "LMM of bay laurel density residuals")
# plot(umca_density@frame$twi15m, resid(umca_density),
#      xlab = "wetness index",
#      ylab = "model residuals",
#      main = "Bay laurel density model residuals vs. wetness index predictor")

**Residual plots of symptomatic leaf count model**
```{r lfct.mod plot, echo=F}

# diagnostic plots
plot_model(lfct.mod, type = "diag")
plot_model(lfct.mod, type = "resid")

# estimate plots
plot_model(lfct.mod, show.values = T, type = "est", 
           title = "Coefficient estimates - leaf count")
plot_model(lfct.mod, show.values = T, type = "std", 
           title = "Standardized coefficient estimates - leaf count")
plot_model(lfct.mod, show.values = T, type = "re")

# effect plots
plot_model(lfct.mod, type = "slope") +
      theme_bw() +
      ylab("log (leaf count + 1)") +
      theme(strip.background = element_blank()) +
      NULL

plot_model(lfct.mod, type = "pred", terms = c("hrs25", "tot_bay.ln1")) +
      theme_classic() +
      ylab("log( leaf count + 1 )") +
      xlab("Hours >25 ºC") +
      ggtitle("Predicted values") +
      NULL

slc_vs_bay_plot <- ggplot(oaks.data, aes(tot_bay.ln1, lfct)) +
      geom_point(alpha = .7) +
      geom_smooth(method = "lm", color = "black") +
      xlab("log( Bay laurel count )") +
      ylab("log( Symptomatic leaf count )") +
      theme_classic()
slc_vs_hrs25_plot <- ggplot(oaks.data, aes(hrs25, lfct)) +
      geom_point(alpha = .7) +
      geom_smooth(method = "lm", color = "black") +
      xlab("Number of hours >25 ºC") +
      ylab("log( Symptomatic leaf count )") +
      theme_classic()

cowplot::plot_grid(slc_vs_bay_plot, slc_vs_hrs25_plot, nrow = 2)

# plot(resid(slc), main = "LMM of symptomatic leaf count residuals")
# qqnorm(resid(slc), main="Q-Q plot for conditional residuals - SLC")
# qqnorm(ranef(slc)$sample_year$`(Intercept)`, 
#        main="Q-Q plot for Sample Year")
# qqnorm(ranef(slc)$plotid$`(Intercept)`, 
#        main="Q-Q plot for plot ID")

# plot(slc@frame$umca_density_ha.log, resid(slc),
#      xlab = "bay laurel density",
#      ylab = "model residuals",
#      main = "SLC model residuals vs. bay laurel density predictor")
# plot(slc@frame$hrs_abv25ds, resid(slc),
#      xlab = "# dry-season hours >25 C",
#      ylab = "model residuals",
#      main = "SLC model residuals vs. dry-season hours >25 C predictor")
# plot(slc@frame$avg_hrs1422_wet_t_t1.log,
#      xlab = "avg. wet-hours 14-22C",
#      ylab = "model residuals",
#      main = "SLC model residuals vs. wet-hours predictor")
# plot(slc@frame$H.2005, resid(slc),
#      xlab = "Shannon's diversity index",
#      ylab = "model residuals",
#      main = "SLC model residuals vs. Shannon's diversity predictor")
# plot(slc@frame$twi15m.log.scl, resid(slc),
#      xlab = "wetness index",
#      ylab = "model residuals",
#      main = "SLC model residuals vs. wetness index predictor")

```{r pairs plot of slc predictor variables, echo=F, eval=F}

pairs(na.omit(select(oakplots.sub, umca_density_ha.log, hrs_abv25ds, avg_hrs1422_wet_t_t1.log, H.2005, twi15m)),
      lower.panel = panel.cor, diag.panel = panel.hist,
      main = "Predictor variables for symptomatic leaf count models")

**Residual plots of average number of wet-hours model**
```{r wet.mod plot, echo=F}
# diagnostic plots
plot_model(wet.mod, type = "diag")
plot_model(wet.mod, type = "resid")

# estimate plots
plot_model(wet.mod, show.values = T, type = "est", 
           title = "Coefficient estimates - wet hours")
plot_model(wet.mod, show.values = T, type = "std", 
           title = "Standardized coefficient estimates - wet hours")
plot_model(wet.mod, show.values = T, type = "re")

# effect plots
plot_model(wet.mod, type = "slope") +
      theme_bw() +
      # ylab("log (leaf count + 1)") +
      theme(strip.background = element_blank()) +
      NULL

# plot(wet_hours_1422, main = "LMM of avg. # of hours 14-22 C current & prev. year residuals")
# plot(wet_hours_1422@frame$umca_density_ha.log.scl, resid(wet_hours_1422),
#      xlab = "bay laurel density",
#      ylab = "model residuals",
#      main = "Wet-hours model residuals vs. bay laurel density predictor")
# plot(wet_hours_1422@frame$twi15m, resid(wet_hours_1422),
#      xlab = "bay laurel density",
#      ylab = "model residuals",
#      main = "Wet-hours model residuals vs. bay laurel density predictor")

**Residual plots of dry-season temperature >25C model**
```{r hot.mod plot, echo=F}

# diagnostic plots
plot_model(hot.mod, type = "diag")
plot_model(hot.mod, type = "resid")

# estimate plots
plot_model(hot.mod, show.values = T, type = "est", 
           title = "Coefficient estimates - hours >25 ºC")
plot_model(hot.mod, show.values = T, type = "std", 
           title = "Standardized coefficient estimates - hours >25 ºC")
plot_model(hot.mod, show.values = T, type = "re")

# effect plots
plot_model(hot.mod, type = "slope") +
      # geom_point() +
      theme_classic() +
      ylab("Number of hours >25 ºC") +
      xlab("Topographic index") +
      theme(strip.background = element_blank(),
            strip.text = element_blank()) +
      NULL

ggplot(oaks.data, aes(twi15m, hrs25)) +
      geom_point(alpha = .3) +
      geom_smooth(method = "lm", color = "black") +
      xlab("Topographic index") +
      ylab("Number of hours >25 ºC") +
      theme_classic()

# plot(hrs_abv25, main = "LMM of # of hours >25 C during prev. dry season residuals")
# plot(hrs_abv25@frame$twi15m, resid(hrs_abv25),
#      xlab = "wetness index",
#      ylab = "model residuals",
#      main = "Dry-season hours model residuals vs. wetness index predictor")
# plot(hrs_abv25@frame$umca_density_ha.log.scl, resid(hrs_abv25),
#      xlab = "wetness index",
#      ylab = "model residuals",
#      main = "Dry-season hours model residuals vs. wetness index predictor")

**Plots of oak infection model**
```{r inf.mod plot, echo=F}

plot_model(inf.mod, type = "diag")
plot_model(inf.mod, type = "resid")

cowplot::plot_grid(
      plot_model(inf.mod, type = "est", show.values = T, 
                 title = "Oak infection - est."),
      plot_model(inf.mod, type = "std", show.values = T, 
                 title = "Oak infection std. est."),
      nrow = 2
      )

plot_model(inf.mod, type = "pred", terms = c("lfct")) +
      # geom_point() +
      theme_classic() +
      xlab("log( symptomatic leaf count + 1 )") +
      ylab("Predicted probability for oak infection") +
      ggtitle("") +
      NULL

ggplot(oaks.data, aes(lfct, inf_oak_ct/(inf_oak_ct+uninf_oak_ct))) +
      geom_point() +
      geom_smooth(method = "loess", color = "black") +
      theme_classic() +
      NULL

# plot(oak_infection, main = "Binomial GLMM of oak infection residuals")
# plot(oak_infection@frame$tot_lfct.log, resid(oak_infection),
#      xlab = "symptomatic leaf count",
#      ylab = "model residuals",
#      main = "Oak infection model residuals vs. symptomatic leaf count predictor")
# plot(oak_infection@frame$avg_hrs1422_wet_t_t1.log, resid(oak_infection),
#      xlab = "avg. wet-hours 14-22C",
#      ylab = "model residuals",
#      main = "Oak infection model residuals vs. wet-hours predictor")
# plot(oak_infection@frame$H.2005, resid(oak_infection),
#      xlab = "Shannon's diversity index",
#      ylab = "model residuals",
#      main = "Oak infection model residuals vs. diversity predictor")

```{r pairs plot of oak infection predictor variables, echo=F, eval=F}

pairs(na.omit(select(oakplots.sub, tot_lfct.log, avg_hrs1422_wet_t_t1.log, H.2005)),
      lower.panel = panel.cor, diag.panel = panel.hist,
      main = "Predictor variables for oak infection model")

## Check for spatial autocorrelation
Check oak infection and symptomatic leaf count response variables and model residuals.

**Oak infection autocorrelation**
```{r spatial autocorrelation inf.mod}

plots_coords <- select(plots_shp@data, PLOT_ID, X, Y)
names(plots_coords) <- tolower(names(plots_coords))
plots_coords <- plots_coords %>% 
      rename(plotid = plot_id) %>% 
      mutate(plotid = tolower(plotid))

oaks.data <- left_join(
      oaks.data, plots_coords
)
summary(oaks.data)

library(spBayes); library(classInt); library(RColorBrewer); library(geoR)
library(ape); library(ncf)

oakplots.coords <- as.matrix(oaks.data[,c("x","y")])
pl.dists <- as.matrix(dist(cbind(oaks.data$x, oaks.data$y)))
pl.dists[1:20,1:20]
dim(pl.dists)
# image(pl.dists)
pl.dists.inv <- 1/pl.dists
# image(pl.dists.inv)
diag(pl.dists.inv) <- 0
pl.dists.inv[is.infinite(pl.dists.inv)] <- 0
pl.dists.inv[1:20,1:20]
# summary(pl.dists.inv)

plot.idist <- iDist(cbind(plots_coords$x, plots_coords$y))
plot.idist[plot.idist==0] <- NA
min(plot.idist, na.rm = T)
max.dist <- .33*max(iDist(cbind(plots_coords$x, plots_coords$y)))

oaks.data <- oaks.data %>% 
      mutate(p.oak.inf = inf_oak_ct/(inf_oak_ct+uninf_oak_ct))
summary(oaks.data)
# ggplot(oakplots.model.data %>% group_by(sample_year), 
#        aes(sample_year, p.oak.inf)) +
#       geom_point(position = "jitter")

Moran.I(oaks.data$p.oak.inf, pl.dists.inv)
cgram.inf <- correlog(oaks.data$x, oaks.data$y, oaks.data$p.oak.inf, increment = 300, resamp = 99)
plot(cgram.inf, main = "Oak infection response correlogram")
# abline(h = mean(cgram.inf$correlation))
max(cgram.inf$correlation)

vario.inf <- variog(coords = oakplots.coords,
                            data = oaks.data$p.oak.inf)
plot(vario.inf)
fit.vario.inf <- variofit(vario.inf)
lines(fit.vario.inf)
abline(h = fit.vario.inf$nugget, col = "purple")
abline(h = fit.vario.inf$cov.pars[1], col = "darkblue")
abline(v = -log(.05)*fit.vario.inf$cov.pars[2], col = "tan4")

infmod.resids <- resid(infmod)
Moran.I(infmod.resids, pl.dists.inv)
cgram.inf.resids <- correlog(oaks.data$x, oaks.data$y, infmod.resids, increment = 300, resamp = 99)
plot(cgram.inf.resids, main = "Oak infection residuals correlogram")

vario.inf.resids <- variog(coords = oakplots.coords,
                            data = infmod.resids)
plot(vario.inf.resids)
fit.vario.inf.resids <- variofit(vario.inf.resids)
lines(fit.vario.inf.resids)
abline(h = fit.vario.inf.resids$nugget, col = "purple")
abline(h = fit.vario.inf.resids$cov.pars[1], col = "darkblue")
abline( v = -log(.05)*fit.vario.inf.resids$cov.pars[2], col = "tan4")
```

There was minimal spatial autocorrelation in the response variable. The spatial autocorrelation that was present was dramitically reduced in the model residuals.

**Symptomatic leaf count**
```{r spatial autocorrelation lfct.mod}
# lfct.data <- oakplots.model.data %>% 
#       select(sample_year, tot_lfct.log) %>% 
#       arrange(sample_year)
# acf(lfct.data$tot_lfct.log)

# qplot(x = sample_year, y = tot_lfct, data = oakplots.model.data)

Moran.I(oaks.data$lfct, pl.dists.inv)
cgram.lfct <- correlog(oaks.data$x, oaks.data$y, oaks.data$tot_lfct.log, increment = 300, resamp = 99)
plot(cgram.lfct, main = "Leaf count response correlogram")

r.lfctmod <- resid(lfctmod)
Moran.I(r.lfctmod, pl.dists.inv)
cgram.lfct.resids <- correlog(oaks.data$x, oaks.data$y, r.lfctmod, increment = 300, resamp = 99)
plot(cgram.lfct.resids, main = "Leaf count residuals correlogram")

There was minimal spatial autocorrelation in the response variable. The spatial autocorrelation that was present was dramitically reduced in the model residuals.

@@

```{r tables of path coefficients, echo=F}

kable(raw_coeff_update, digits = 3,
      caption = "Raw path coefficients")
# kable(raw_coeff_refit, digits = 3,
#       caption = "Raw path coefficients")
# kable(raw_coefficients_noUMCA,
#       caption = "Raw path coefficients - no UMCA density")
